---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# detectnorm

<!-- badges: start -->
<!-- badges: end -->

The goal of detectnorm is to provide an eduacated guess on nonnormality without raw data based on the Beta distribution. 
When conducting a meta-analysis for two independent groups, we generally retrieved very limited statistics from primary studies, and it is hard to access the raw data. However, the non-normality of raw data could influence the meta-analytic results greatly (Sun & Cheung, 2020).
his package allows meta-analysis researchers to speculate the skewness and kurtosis of the raw data, even without the raw data. 

The package provides not only the skewness and kurtosis estimates but also the figures to visualize them. Now, this package could work directly with the standardized mean difference for two independent and dependent groups. It also works for one group of data.

It makes it easier to have a quick check for meta-analysis with only means, standard deviations, and possible ranges (not sample minimum and sample maximum). We also provide a check for the tails directions of the two groups for meta-analysis. When the two tails are pointing towards each other (check_tail: "toward"), together with the skewnesses > 1.0 or < -1.0, researchers should be carefully use the normal way for meta-analytic analyses. 

## Documentation
A good start to understand the problems of non-normality in primary studies on meta-analytic results is the following paper:

Sun, R. W., & Cheung, S. F. (2020). The influence of nonnormality from primary studies on the standardized mean difference in meta-analysis. *Behavior Research Methods*, 52(4), 1552-1567.

## Installation

You can install the official version within R with (not the most updated version): 

```r
install.packages("detectnorm")
```

You can install the development version of detectnorm from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("irissun/detectnorm")
```

## Example

This is a basic example which shows you how to solve a common problem:

For one study if you assumed non-normal distribution in population:

```{r example}
devtools::load_all()
#library(detectnorm)
#nonnormal 
set.seed(65180)
#Using Fleishman's method to generate non-normal data
dat1 <- rnonnorm(n = 1000, mean = 0, sd = 1, skew = 2, kurt = 5)$dat
psych::describe(dat1)
hist(dat1)
```

Suppose we don't have the raw data. We could only get the means, standard deviations, ranges from one paper. 

```{r desbeta}
ex1 <- desbeta(vmean = mean(dat1), vsd = sd(dat1), lo = min(dat1), 
               hi = max(dat1), showFigure = TRUE)
ex1
```

The skewness `ex1$bskew` and kurtosis `ex1$bkurt` estimated from beta distribution are `r round(ex1$bskew,3)` and `r round(ex1$bkurt,3)`. The values are very close to 2 that set at the beginning. 

The function `des2beta` works well for independent two groups and dependent two groups with figure to give direct impression on the nonnormal problems.

```{r}
#independent two groups
des2beta(vmean=c(1.63, 2.41), vsd=c(1.137, .930),hi = c(5, 5), lo = c(0, 0), design = "ind", showFigure = TRUE)
#dependent two groups
des2beta(vmean=c(1.83, 2.41), vsd=c(1.137, .930),hi = c(5, 5), lo = c(0, 0), design = "dep", cor_xy = .5, showFigure = TRUE)
```

For one meta-analysis, which may have studies that contain dependent two groups and independent two groups, the function works well for the dataset used for package `metafor`. You need to add columns for `hi1` (i.e., maximum value for group 1), `hi2` (i.e., maximum value for group 2), `lo1` (i.e., minimum value for group 1), `lo2` (i.e., minimum value for group 2), `design` (i.e., research design: either "dep" or "ind"), and `cor_xy` (i.e., correlations for dependent two groups, `NA` for independent two groups). 
